Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2015-01-15 paper/sheaf_lics.tex
in the middle of section 4

\subsection{Lawvere-Tierney Sheafification in Topos Theory}

@@

The Giraud-Rezk-Lurie axioms are the $\infty$-version of Giraud's
axioms that characterize a topos. Namely, there are 4 axioms on a
$(\infty,1)$-category that have been shown to be equivalent to be
equivalent to $(\infty,1)$-topos~\cite[Chapter 6]{lurie}. This is not
the purpose of this paper to discuss those axioms, we only state here
a particular case of the last axiom which says that every groupoid
object is effective. 
%
This particular case connects an epimorphism (or $-1$-truncated
function) to the colimit of its \v{C}ech nerve. 
%
This is the $\infty$-generalization of the fact that, in a topos,
every epimorphism is the colimit of its kernel pair, which plays a
central part in the definition of sheafification.

@@

Moreover the following diagram commute
$$\xymatrix{
  \sum_{x:X} (f x = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f x = y)\right) \ar@{=}[dl]\\
  \sum_{x:\modal X} (\modal f x = \eta_Y y) &
}$$
where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and $\pi_2 \circ
\gamma$ is given by the modalisation of paths.
\end{prop}
\begin{proof}
\nt{ça a du sens de mettre une preuve, non ? c'est non trivial et
  nouveau par rapport aux travaux sur les modalités.}
\end{proof}

@@

We will define sheaves by induction on the level of homotopy type:
the base case will be any left exact modality on $\HProp$, and we
give in this section some definitions useful for the inductive case.

@@

In this section, we suppose given a truncation index $n\geqslant -1$,
and a left exact modality $\modal$ on $\Type_n$.

\begin{defi}
  Let $E$ be a type. The {\em closure} of a subobject of $E$ classified by $\chi$
  is the subobject of $E$ classified by $\modal \circ \chi$.

  The subobject of $E$ classified by $\chi$ is said {\em closed in
    $E$} if its closure is itself : $\chi = \modal \circ \chi$.
\end{defi}

\begin{defi}
  Let $E$ be a type, and $\chi:E \to \Type_n$. The subobject $A$ of $E$
  classified by $\chi$ is {\em dense} in $E$ if its $\modal$-closure
  is $E$ seen as a subobject of $E$, \ie
  $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq (\modal (\chi~e)).$$ 
  Moreover, for any $x:A$ we need the following coherence diagram to
  commute
  $$\xymatrix{
    \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_\iota & (\chi~x) \ar[d]^{\eta_{(\chi~x)}}\\
    \sum_{e':E} x = e' \ar@{=}[r] & \modal (\chi~x)
  }$$
  where $\iota: x \mapsto ({x_1}_1 ; x_2)$.
\end{defi}

@@

\begin{defi}[Restriction]
  For any $A, E:\Type$ and $F:\Type_{n+1}$, we define~:
  \begin{itemize}
  \item if $A \hookrightarrow E$, $M_E^\chi$ is the map sending an
    arrow $f~:~E\to F$ to $\restriction f A$ the restriction of $f$ to $A$.
  \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
    map sending $f~:~E \to f$ to $\restriction f A$.
  \end{itemize}
\end{defi}

>>>>

\subsection{Lawvere-Tierney Sheafification in Topos Theory}
\label{sec:lawv-tiern-sheaf}

@@

The Giraud-Rezk-Lurie axioms are the $\infty$-version of Giraud's
axioms that characterize a topos. Namely, there are 4 axioms on a
$(\infty,1)$-category that have been shown to be equivalent to be
equivalent to $(\infty,1)$-topos~\cite[Chapter 6]{lurie}. This is not
the purpose of this paper to discuss those axioms, we only state here
a particular case of the last axiom which says that every groupoid
object is effective. 
%
This particular case connects an epimorphism (or a $-1$-connected
function) to the colimit of its \v{C}ech nerve.
%
This is the $\infty$-generalization of the fact that, in a topos,
every epimorphism is the colimit of its kernel pair, which plays a
central part in the definition of sheafification.

@@

Moreover the following diagram commute
$$\xymatrix{
  \sum_{x:X} (f x = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f x = y)\right) \ar@{=}[dl]\\
  \sum_{x:\modal X} (\modal f x = \eta_Y y) &
}$$
where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and $\pi_2 \circ
\gamma$ is given by the modalisation of paths.
\end{prop}
\begin{proof}
\nt{ça a du sens de mettre une preuve, non ? c'est non trivial et
  nouveau par rapport aux travaux sur les modalités. Ou alors c'est
  la preuve de Shulman qui tu n'as pas faite. En tout cas, il faut clarifier.}
\end{proof}

@@

As explained in the introduction, a Lawvere-Tierney topology for an
topos $\E$, as defined in Section~\ref{sec:lawv-tiern-sheaf},
corresponds to a left exact modality at level $-1$.
%
And sheafification corresponds to lifting this modality to a left
exact modality at level $0$.
%
In homotopy type theory, it is possible to use similar ideas to lift
any left exact modality at level $n$ to a left exact modality at level
$n+1$ using sheafification.

In this section, we suppose given a truncation index $n\geqslant -1$,
and a left exact modality $\modal$ on $\Type_n$ and we introduce the
definition of sheaves for $(n+1)$-types and basic properties of
sheaves that will be useful in the next section.

% We will define sheaves by induction on the level of homotopy type:
% the base case will be any left exact modality on $\HProp$, and we
% give in this section some definitions useful for the inductive case.

@@

To define sheaves, we first have to leverage the notion of dense
subobjects for the modality $\modal$. The closure operator is defined
using the $n$-subobject classifier $\Type_n$ introduced in
Section~\ref{ssec:hott}.

\begin{defi}
  Let $E$ be a type. The {\em closure} of a subobject of $E$ with
  n-truncated homotopy fibers (or $n$-subobject of $E$, for short),
  classified by $\chi : E \to \Type_n$, is the subobject of $E$
  classified by $\modal \circ \chi$.

  An $n$-subobject of $E$ classified by $\chi$ is said to be {\em
    closed in $E$} if it is equal to its closure, \ie
  $\chi = \modal \circ \chi$.
\end{defi}

The definition of a dense $n$-subobject requires more work as the
characteristic function $\chi_E$ of $E$ seen as an $n$-subobject of
itself is not the constant function equal to $\one$ as it is the case
for topoi, but rather the map 
%
$$\chi_E  = \lambda e . \ \sum_{e':E} e=e'.$$
%
This means that we need to take more care of the homotopy structure of
this charactestic function, and in particular, the defintiion of a
dense $n$-subobject comes with an additional coherence condition
(which is trivial when $n=-1$).

\begin{defi}
  Let $E$ be a type, and $\chi:E \to \Type_n$. The subobject $\iota :
  A \to E$ of $E$
  classified by $\chi$ is {\em dense} in $E$ when its $\modal$-closure
  is equvialent to $\chi_E$, \ie
  %
  $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq (\modal
  (\chi~e)).$$
  % 
  and moreover, for any $x:A$, the following diagram 
  commutes, 
  $$\xymatrix{
    \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_{\iota'} & (\chi~x) \ar[d]^{\eta_{(\chi~x)}}\\
    \sum_{e':E} x = e' \ar@{=}[r] & \modal (\chi~x)
  }$$
  where $\iota': x \mapsto (\iota (\pi_1 x) ; \pi_2 x)$.
\end{defi}

@@

As recall in Section~\ref{sec:lawv-tiern-sheaf}, in topos theory,
sheaves are elements that can not distinguish between objects and
their dense sub-objects. Here, we need to distinguish between
dense $(-1)$-subobjects, that will be used in the definition of
sheaves, and dense $n$-subobjects, that will be used in the definition
of separated types.

\begin{defi}
  For any $A, E:\Type$ and $F:\Type_{n+1}$, we define~:
  \begin{itemize}
  \item if $m : A \hookrightarrow E$ is a monomorphism, we define 
    %
    $$
    M_E^\chi : (E \to F) \to (A \to F)
    $$
    %
    is the map sending an
    arrow $f~:~E\to F$ to $\restriction f A$ the restriction of $f$ to $A$.
  \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
    map sending $f~:~E \to f$ to $\restriction f A$.
  \end{itemize}
\end{defi}
